---
title: "E.elymi_DEseq2_STR_INF"
author: "Kate"
date: "3 April 2019"
output: html_document
---
---
title: "E.elymi_DEseq2_PS_STR"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


# DESeq2 installation
#if (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
#BiocManager::install("DESeq2", version = "3.8")

# tximport installation
#if (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
#BiocManager::install("tximport", version = "3.8")

# tximportData installation to see example data
#if (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
#BiocManager::install("tximportData", version = "3.8")

#install.packages("readr")

#install.packages("tidyverse")


library("DESeq2")
library("tximport")
library("tximportData")
library("readr")
library("tidyverse")
```

Following example from http://127.0.0.1:11302/library/DESeq2/doc/DESeq2.html#transcript-abundance-files-and-tximport-input

```{r get_data}

# import sample list and assign column and row names
samples <- read.table("../../config.txt", header = FALSE)
colnames(samples) <- c("pop", "center", "assay", "run_number", "lane", "sample", "experiment", "run", "condition")
rownames(samples) <- samples$sample

# subset only one species
samples <- samples[samples$pop == "E.elymi_NfE728",]
samples

# filter out PS
samples <- samples[samples$condition != "PS",]
samples
```


```{r example_cont}
dir <- "../../2_salmon"
# use external HD files
dir <- "/mnt/usb-Seagate_Expansion_NAA7H6EA-0:0-part2/Epichloe/2_salmon"
files <- file.path(dir,"salmon_quant", samples$pop, samples$sample, "quant.sf")
names(files) <- samples$sample

tx2gene <- read.table(file.path(dir, "transcriptome/E.elymi_E728/transcriptome.headers.txt"))
colnames(tx2gene) <- c("TXNAME", "GENEID")
tx2gene
```


```{r}
txi <- tximport(files, type="salmon", tx2gene=tx2gene)
```


```{r}
library("DESeq2")
ddsTxi <- DESeqDataSetFromTximport(txi,
                                   colData = samples,
                                   design = ~ condition)
```

DESeq2 Analysis

filter out data with less than 10 read counts
```{r filter}
keep <- rowSums(counts(ddsTxi)) >= 10
ddsTxi <- ddsTxi[keep,]


```

if you never tell the DESeq2 functions which level you want to compare against (e.g. which level represents the control group), the comparisons will be based on the alphabetical order of the levels. Here just set the reference condition
```{r set_ref_condition}
ddsTxi$condition <- relevel(ddsTxi$condition, ref = "INF")
```

```{r DESeq}
dds <- DESeq(ddsTxi)
#colData(dds)
res <- results(dds)
res
```


```{r diff_gene_exp}
summary(res) #summary of results
```


```{r summary_sorted_by_p}

res <- res[order(res$padj),]
head(res)
```


```{r plot_top_siz_diffs}

par(mfrow=c(2,3))
plotCounts(dds, gene="FUN_000621", intgroup="condition")
plotCounts(dds, gene="FUN_001518", intgroup="condition")
plotCounts(dds, gene="FUN_001928", intgroup="condition")
plotCounts(dds, gene="FUN_003127", intgroup="condition")
plotCounts(dds, gene="FUN_003281", intgroup="condition")
plotCounts(dds, gene="FUN_003546", intgroup="condition")
```


```{r }

#reset par
par(mfrow=c(1,1))
# Make a basic volcano plot
with(res, plot(log2FoldChange, -log10(pvalue), pch=20, main="Volcano plot", xlim=c(-3,3)))

# Add colored points: blue if padj<0.01, red if log2FC>1 and padj<0.05)
with(subset(res, padj<.01 ), points(log2FoldChange, -log10(pvalue), pch=20, col="blue"))
with(subset(res, padj<.01 & abs(log2FoldChange)>2), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
```

```{r PCA}
#First we need to transform the raw count data
#vst function will perform variance stabilizing transformation

vsdata <- vst(dds, blind=FALSE)

plotPCA(vsdata, intgroup="condition") #using the DESEQ2 plotPCA fxn we can
```


Get data from gff3 file
ID genes that are over 2 fold change that are close to one another

```{r ID_clusters}

positions <- read.table("../../0_raw_data/genome_assembly/Eel_728/Epichloe_elymi.gff3", header = FALSE, comment.char = "#", blank.lines.skip = TRUE, fill = TRUE, sep = "\t")
colnames(positions) <- c("seqid", "source", "type", "start", "end", "score", "strand", "phase", "attributes")

# select out mRNA (these are the ones with product described)
mRNA <- positions[positions$type == "mRNA",]

# split attributes to get c("trans_id", "gene_id", "product")
temp <- str_split(mRNA$attributes, ";", n=3)
df <- as.data.frame(t(as.data.frame(temp)))
rownames(df)<-NULL
mRNA[, c("trans_id", "gene_id", "product")] <- df
mRNA$trans_id <- str_replace(mRNA$trans_id, "ID=", "")
mRNA$gene_id <- str_replace(mRNA$gene_id, "Parent=", "")
mRNA$product <- str_replace(mRNA$product, "product=", "")
mRNA

# subset target results that have an adjusted pvalue less than .01 and a log2FoldChange >2
targets <- rownames(subset(res, padj<.01 & abs(log2FoldChange)>2))
targets


GOI <- mRNA[mRNA$gene_id %in% targets,]
table <- GOI %>% group_by(seqid, start)
table
write.table(table, "E.elymi_INF_STR_log2foldchange.txt", quote = FALSE, row.names = FALSE, sep = "\t")

```